---
title: "CBAM’s Effect on Great Britain’s Electricity Flows"
author: "Luca Taschini, Gian Luca Vriz, Joshua Burke"
date: "2024-07-08"
output:
  pdf_document: default
  html_document: default
---

# Libraries

The libraries utilized for the project are delineated below.

```{r}
library(gtools)
library(cowplot)
library(plyr)
library(viridis)
library(FinTS)
library(lubridate)
library(data.table)
library(zoo)
library(forecast)
library(lmtest)
library(stringr)
library(zoo)
library(corrplot)
library(readxl)
library(readr)
library(countrycode)
library(ggplot2)
library(tidyr)
library(dplyr)
```

# Data

Initially, electricity prices are uploaded for both the Great Britain (GB) and the European Union (EU).

```{r}
#EPEX spot day ahead UK baseload
Electricity_price_gb <-
  read_csv("Data/Electricity_price_UK_EPEX2.csv", skip = 1)
colnames(Electricity_price_gb) <- c('Date', 'Daily_ave_GB')
Electricity_price_gb$Date <- dmy(Electricity_price_gb$Date)
Electricity_price_gb$Daily_ave_GB <-
  as.numeric(Electricity_price_gb$Daily_ave_GB)
#Period
Electricity_price_gb <-
  Electricity_price_gb[Electricity_price_gb$Date < '2024-01-01' &
                         Electricity_price_gb$Date > '2019-12-31', ]
```

For the EU's electricity price, the countries with the highest electricity trade volumes with the GB are considered. These include:

* France,
* Belgium,
* Denmark,
* Netherlands,
* Ireland.

```{r}
#Electricity price EU
#Raw data daily values
#https://ember-climate.org/data-catalogue/european-wholesale-electricity-price-data/
Electricity_price_EU <-
  read.csv("Data/Electricity_price_EU.csv")
Electricity_price_EU <-
  Electricity_price_EU[Electricity_price_EU$Country == "Belgium" |
                         Electricity_price_EU$Country == "Ireland" |
                         Electricity_price_EU$Country == "Denmark" |
                         Electricity_price_EU$Country == "France" |
                         Electricity_price_EU$Country == "Netherlands", ]
Electricity_price_EU$Date <- as.Date(Electricity_price_EU$Date)
#Average
Electricity_price_EUg <- Electricity_price_EU  %>%
  group_by(Date) %>%
  dplyr::summarise(Average.price = mean(Price..EUR.MWhe.))
colnames(Electricity_price_EUg) <- c('Date', 'Daily_ave_EU')
#Period
Electricity_price_EUg <-
  Electricity_price_EUg[Electricity_price_EUg$Date < '2024-01-01' &
                          Electricity_price_EUg$Date > '2019-12-31', ]
str(Electricity_price_EUg)
```

To convert currency between the two regions, the $€/£$ exchange rate is adopted.

```{r}
#Exchange rate
#Raw data
#https://www.investing.com/currencies/eur-gbp-historical-data
EUR_GBP <-
  read.csv("Data/EUR_GBP_change_data.csv")
EUR_GBP$Date <- dmy(EUR_GBP$Date)
EUR_GBP <- EUR_GBP[, 1:2]
str(EUR_GBP)
```

Finally, the overall dataset is merged on a daily basis. Additionally, the difference between the GB electricity price and the EU electricity price is calculated (using log-transformations, i.e. GB/EU).

```{r}
#Merging data
merged_data <- merge(EUR_GBP, Electricity_price_EUg, by = 'Date')
merged_data <- merge(Electricity_price_gb, merged_data, by = 'Date')
#From EU to £
merged_data$Daily_ave_EU <-
  merged_data$Daily_ave_EU * merged_data$Price
merged_data$Monthly.Date <-
  as.Date(as.yearmon(format(merged_data$Date, "%Y-%m")))
#Only positive values
merged_data <- merged_data[merged_data$Daily_ave_GB > 0 &
                             merged_data$Daily_ave_EU > 0, ]
#log
merged_data$Daily_ave_GB <- log(merged_data$Daily_ave_GB)
merged_data$Daily_ave_EU <- log(merged_data$Daily_ave_EU)
#Diff
merged_data$Price_diff <-
  (merged_data$Daily_ave_GB - merged_data$Daily_ave_EU)
#NA checking
merged_data <- na.omit(merged_data)
```

# Electricity prices and Great Britain’s net imports

```{r}
#Data
# gco2/KW -> carbon intensity
# MW -> electricity imports/exports/production/consumption
# £/MW -> EU/UK price
# £/tco2 -> EU/UK ETS
```


Data from Electricity Map is utilized to conduct an empirical analysis on the relationship between GB-EU electricity prices and GB's net imports.

```{r,warning=FALSE}
#Uploading electricity map data
GB_20 <-
  read.csv("C:/Users/gian-/Desktop/Electricity_CBAM/Data/Electricity_map_GB/GB_20.csv")[, 2:82]
GB_21 <-
  read.csv("C:/Users/gian-/Desktop/Electricity_CBAM/Data/Electricity_map_GB/GB_21_22.csv")[, 2:62]
GB_23 <-
  read.csv("C:/Users/gian-/Desktop/Electricity_CBAM/Data/Electricity_map_GB/GB_23.csv")[, 2:62]
dataset_tot_GB <- bind_rows(GB_20, GB_21, GB_23)
dataset_tot_GB[is.na(dataset_tot_GB)] <- 0
#Net imports EU
dataset_tot_GB$Net_imports_EU <-
  dataset_tot_GB$power_net_import_NL_avg +
  dataset_tot_GB$power_net_import_IE_avg + dataset_tot_GB$power_net_import_FR_avg +
  dataset_tot_GB$power_net_import_BE_avg + dataset_tot_GB$power_net_import_DK.DK1_avg
dataset_tot_GB <- dataset_tot_GB[, -c(2:5, 17, 62:81)]
dataset_tot_GB$datetime <- as.Date(date(dataset_tot_GB$datetime))
#Co2 (gCo2) total
dataset_tot_GB$CO2_production_avg <-
  dataset_tot_GB$carbon_intensity_production_avg *
  dataset_tot_GB$total_production_avg * 1000 #Form MW to gco2 by gco2/KW
#Co2 (gCo2) direct
dataset_tot_GB$CO2_production_direct_avg <-
  dataset_tot_GB$carbon_intensity_direct_avg *
  dataset_tot_GB$total_production_avg * 1000
```

All variables on the platform are expressed on an hourly basis. Therefore, a daily aggregation is performed for a selected subset of variables.

```{r,warning=FALSE}
#Daily aggregation
dataset_GB_daily <- dataset_tot_GB[, c(
  "datetime",
  "CO2_production_avg",
  "CO2_production_direct_avg",
  "total_production_avg",
  "total_consumption_avg",
  "total_import_avg",
  "total_export_avg",
  "Net_imports_EU"
)]
dataset_GB_daily <- dataset_GB_daily %>% group_by(datetime) %>%
  dplyr::summarize(across(everything(), sum, na.rm = TRUE))
colnames(dataset_GB_daily)[1] <- 'Date'
```

Subsequently, the dynamics of GB's trades with the world are plotted on an annual basis.

```{r,fig.dim = c(15,9)}
dataset_GB_daily$Year <- year(dataset_GB_daily$Date)
plot_data <-
  dataset_GB_daily[, c("Year", "total_import_avg", "total_export_avg")] %>%
  group_by(Year) %>%
  dplyr::summarize(across(everything(), sum, na.rm = TRUE))
plot_data$total_import_avg <-
  plot_data$total_import_avg / 1000 #From MW to GW
plot_data$total_export_avg <-
  plot_data$total_export_avg / 1000 #From MW to GW
colnames(plot_data) <- c('Year', 'Imports', 'Exports')
meltdata <- melt(plot_data, id = c('Year'))
colnames(meltdata) <- c('Year', 'Trade', 'Value')
meltdata %>%
  ggplot(aes(
    x = Year,
    y = Value,
    group = Trade,
    fill = Trade
  )) +
  geom_col(
    position = position_dodge(width = 0.6),
    colour = "black",
    width = 0.6
  ) +
  geom_hline(aes(yintercept = 0)) + ylab('GW') + ggtitle('GB Electricity Trade') +
  scale_colour_manual(values = c('blue', 'red'),
                      aesthetics = c("colour", "fill"))
```
The same is done considering the dynamics of GB's trades with the EU only (data from GOV-UK).

```{r}
#https://www.gov.uk/government/collections/energy-trends#2024-data
#Raw data quarterly
Electricity_imp_exp <-
  read_excel("Data/Electricity_imp_exp.xlsx",
             sheet = "Quarterly imports and exports",
             skip = 5)
#GB imports/exports/Net imports to EU.
Electricity_imp_exp <-
  Electricity_imp_exp[, c(1, 3, 4, 9, 10, 21, 22)]
colnames(Electricity_imp_exp) <- c(
  'Year',
  'Imports_GB',
  'Exports_GB',
  'Imports_NIR',
  'Exports_NIR',
  'Imports_NR',
  'Exports_NR'
)
#Filtering
Electricity_imp_exp$Imports_NR <-
  as.numeric(Electricity_imp_exp$Imports_NR)
Electricity_imp_exp$Exports_NR <-
  as.numeric(Electricity_imp_exp$Exports_NR)
Electricity_imp_exp[is.na(Electricity_imp_exp)] <- 0
Electricity_imp_exp$Imports_GB <-
  Electricity_imp_exp$Imports_GB - Electricity_imp_exp$Imports_NIR -
  Electricity_imp_exp$Imports_NR
Electricity_imp_exp$Exports_GB <-
  Electricity_imp_exp$Exports_GB - Electricity_imp_exp$Exports_NIR -
  Electricity_imp_exp$Exports_NR
#Rearranging
Electricity_imp_exp <- Electricity_imp_exp[, c(1:3)]
colnames(Electricity_imp_exp) <-
  c('Year', 'Imports', 'Exports')
Electricity_imp_exp <-
  Electricity_imp_exp[Electricity_imp_exp$Year > 2019, ]
Electricity_imp_exp$Imports <-
  as.numeric(Electricity_imp_exp$Imports)
Electricity_imp_exp$Exports <-
  as.numeric(Electricity_imp_exp$Exports)
Electricity_imp_exp[is.na(Electricity_imp_exp)] <- 0
```

```{r,fig.dim = c(15,9)}
Electricity_imp_exp <- Electricity_imp_exp %>%
  group_by(Year) %>%
  summarize(across(everything(), sum, na.rm = TRUE))
meltdata <- melt(Electricity_imp_exp, id = c('Year'))
colnames(meltdata) <- c('Year', 'Trade', 'Value')
meltdata %>%
  ggplot(aes(
    x = Year,
    y = Value,
    group = Trade,
    fill = Trade
  )) +
  geom_col(
    position = position_dodge(width = 0.6),
    colour = "black",
    width = 0.6
  ) +
  geom_hline(aes(yintercept = 0)) + ylab('GW') + ggtitle('Great Britain\'s Electricity Trade with the European Union') +
  scale_colour_manual(values = c('blue', 'red'),
                      aesthetics = c("colour", "fill"))
```

The following chunk provides a descriptive summary of the daily values, focusing on the differences in prices between Great Britain and the European Union, as well as net imports.

```{r,warning=FALSE,fig.dim = c(15,9)}
#Merging -> observations loss is due to finance data not reported for some days
dataset_GB_daily <-
  merge(merged_data[, c(1, 2, 4, 6)], dataset_GB_daily, by = 'Date')
#Total exports
dataset_GB_daily$Net_imports_tot <-
  dataset_GB_daily$total_import_avg -
  dataset_GB_daily$total_export_avg
#Carbon intensity production (all)
dataset_GB_daily$carbon_intensity_production_avg <-
  dataset_GB_daily$CO2_production_avg /
  (dataset_GB_daily$total_production_avg * 1000) #gco2/KW
#Carbon intensity production (direct)
dataset_GB_daily$carbon_intensity_production_direct_avg <-
  dataset_GB_daily$CO2_production_direct_avg /
  (dataset_GB_daily$total_production_avg * 1000) #gco2/KW
#Plots
data_plot <- dataset_GB_daily[dataset_GB_daily$Date < '2023-01-01' &
                                dataset_GB_daily$Date > '2021-12-31', ]
data_plot <- dataset_tot_GB[dataset_tot_GB$datetime < '2023-01-01' &
                              dataset_tot_GB$datetime > '2021-12-31', ]
A1 <- ggplot(dataset_GB_daily, aes(x = Date, y = Price_diff)) +
  geom_line(aes(x = as.Date('2022-1-1')), linetype = "dashed", col = 'red') +
  geom_line(aes(x = as.Date('2023-1-1')), linetype = "dashed", col = 'red') +
  geom_point() +
  geom_smooth(method = "gam") + xlab('Time') + ylab('Value') + ggtitle('Price difference in log-scale (£/MW)')

A2 <-
  ggplot(dataset_GB_daily, aes(x = Date, y = Net_imports_EU / 1000)) + # from MW to GW
  geom_line(aes(x = as.Date('2022-1-1')), linetype = "dashed", col = 'red') +
  geom_line(aes(x = as.Date('2023-1-1')), linetype = "dashed", col = 'red') +
  geom_point() +
  geom_smooth(method = "gam") + xlab('Time') + ylab('GW') + ggtitle('Power net imports with the EU')

plot_grid(A1, A2, nrow = 1, scale = 1)
```

Finally, a restricted dataset containing all observations from 2022 is used to estimate an ARIMAX model, incorporating GB-EU price differences as an exogenous variable.

```{r,fig.dim = c(15,9)}
#Restricted dataset for simulations
dataset_GB_daily <-
  dataset_GB_daily[dataset_GB_daily$Date > '2022-01-01', ]
#GW
dataset_GB_daily$Net_imports_EU <-
  dataset_GB_daily$Net_imports_EU / 1000 #from MW to GW
#Lag-values for overconfidence in predictions
dataset_GB_daily$Price_difflag <- lag(dataset_GB_daily$Price_diff)
#Arimax
Arimax <- auto.arima(
  dataset_GB_daily$Net_imports_EU,
  xreg = matrix(c(dataset_GB_daily$Price_diff), ncol = 1),
  max.p = 10,
  max.q = 10,
  max.d = 10,
  lambda = NULL
)
#Residuals
checkresiduals(Arimax)
bptest(residuals(Arimax) ~ fitted(Arimax))
#Coefficients
coeftest(Arimax)
```
The residuals of the model appear to be homoscedastic, normally distributed, and free from autocorrelation.

```{r,include = FALSE}
coefficients <- coef(Arimax)
standard_errors <- sqrt(diag(vcov(Arimax)))
p_values <- 2 * (1 - pnorm(abs(coefficients / standard_errors)))
results <- data.frame(Estimate = coefficients,
                      `Std. Error` = standard_errors,
                      `P-Value` = p_values)
results
# Function to create LaTeX table
create_latex_table <- function(results) {
  cat("\\begin{table}[!htbp] \\centering\n")
  cat("\\caption{ARIMAX Model Results}\n")
  cat("\\begin{tabular}{@{\\extracolsep{5pt}}lccc}\n")
  cat("\\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex]\n")
  cat("& Estimate & Std. Error & P-Value \\\\\n")
  cat("\\hline \\\\[-1.8ex]\n")
  
  for (i in 1:nrow(results)) {
    cat(
      rownames(results)[i],
      "&",
      format(results$Estimate[i], digits = 4),
      "&",
      format(results$Std..Error[i], digits = 4),
      "&",
      format(results$P.Value[i], digits = 4),
      "\\\\\n"
    )
  }
  
  cat("\\hline \\\\[-1.8ex]\n")
  cat("\\end{tabular}\n")
  cat("\\end{table}\n")
}

# Call the function
create_latex_table(results)
```

# Scenarios

Finally, a simulation study is conducted to evaluate the possible impact of the EU CBAM. In order to perform the exercise, the CBAM obligation is calculated based on the embedded
carbon emissions of GB direct electricity production. Then, the CBAM adjustment is determined by the difference between the price of carbon already paid in the UK and the price of carbon in Europe.

```{r}
#EUA
EUA_dec <- read_excel("Data/EUA_dec_GBP.xlsx")
EUA_dec$Date <- as.Date(date(EUA_dec$Date))
EUA_dec <- na.omit(EUA_dec)
#UKA
UKA_dec <- read_excel("Data/UKA_dec.xlsx")
UKA_dec$Date <- as.Date(date(UKA_dec$Date))
UKA_dec <- na.omit(UKA_dec)
#Merge
merged_ETS <- merge(UKA_dec, EUA_dec, by = 'Date')
colnames(merged_ETS) <- c('Date', 'UKA', 'EUA')
merged_ETS$UKA <- as.numeric(merged_ETS$UKA)
merged_ETS$EUA <- as.numeric(merged_ETS$EUA)
```

The exercise is conducted on the solar year 2023.

```{r,fig.dim = c(15,9)}
Plot <- melt(merged_ETS, id.vars = c('Date'))
colnames(Plot)[2] <- c('ETS')
B1 <- ggplot(data = Plot, aes(
  x = Date,
  y = value,
  group = ETS,
  colour = ETS
)) + geom_line() + labs(y = "£/tco2", x = "Time") + ggtitle("UK and EU's ETS") +                                                  # Modify labels and colors
  scale_color_manual(
    name = 'ETS',
    labels = c("UKA", "EUA"),
    values = c("red", "blue")
  )

merged_ETS$Diff <- (merged_ETS$EUA - merged_ETS$UKA)
merged_ETS[merged_ETS$Diff <= 0, 'Diff'] <- 0

B2 <- ggplot(data = merged_ETS, aes(
  x = Date,
  y = Diff,
  colour = 'darkgreen'
)) + geom_line() + labs(y = "£/tco2", x = "Time") + ggtitle("Permit Price Differences (+)") +                                                  # Modify labels and colors
  scale_color_manual(
    name = 'Variable',
    labels = c(expression(Delta[ETS])),
    values = c("darkgreen")
  )

plot_grid(B1, B2, nrow = 2, scale = 1)
```

```{r}
merged_ETS <- merged_ETS[merged_ETS$Date > '2022-12-31' &
                           merged_ETS$Date < '2024-01-01',]
```

Due to the temporal characteristics of the ARIMAX model, both the real scenario and the projected CBAM scenario are filtered through the baseline model, which assumes no differences in electricity prices. 

```{r}
dataset_scenario <-
  dataset_GB_daily[dataset_GB_daily$Date > '2022-12-31' &
                     dataset_GB_daily$Date < '2024-01-01', ]
#From log to baseline
dataset_scenario$Daily_ave_GB <- exp(dataset_scenario$Daily_ave_GB)
dataset_scenario$Daily_ave_EU <- exp(dataset_scenario$Daily_ave_EU)
dataset_scenario <- merge(dataset_scenario, merged_ETS, by = 'Date')
#CBAM
pred <- forecast(Arimax, xreg = c(rep(0, 251)), h = 251)
xreg_CBAM <- log((dataset_scenario$Daily_ave_GB + (
  (
    dataset_scenario$carbon_intensity_production_direct_avg / 1000
  ) * dataset_scenario$Diff
)) / dataset_scenario$Daily_ave_EU) #/1000 to convert gco2/KW into tco2/MW
pred1 <- forecast(Arimax, xreg = c(xreg_CBAM), h = 251)
#Real
xreg_real <-
  log(dataset_scenario$Daily_ave_GB / dataset_scenario$Daily_ave_EU)
pred2 <- forecast(Arimax, xreg = c(xreg_real), h = 251)
```

The results of the exercise indicate that the introduction of the EU Carbon Border Adjustment Mechanism (CBAM) could lead to a reduction in net electricity exports to the EU by approximately 0.45 GW per day.

```{r,fig.dim = c(15,9)}
#Results
meltData <- melt(data.frame(
  CBAM = mean(pred1$mean - pred$mean),
  Real = mean(pred2$mean - pred$mean)
))
#Yearly
meltData$value <- meltData$value * 365
ggplot(meltData, aes(
  x = reorder(variable, value),
  y = value,
  fill = variable,
  color
)) +
  geom_bar(position = "dodge",
           stat = "summary",
           fun = "identity") +
  theme_bw(base_size = 14) +
  scale_fill_manual(values = c('blue', 'red')) +
  scale_alpha_manual(values = c(1, 0.1)) +
  theme(legend.position = "none") +
  xlab("Scenario") + ggtitle('Net imports, GB', 'Baseline, model with no difference in prices') +
  ylab('GW') + scale_x_discrete(labels = c('Historical', 'CBAM'))
#-166 GW yearly
#-0.45 GW daily
#Mean price change
mean(
  dataset_scenario$carbon_intensity_production_direct_avg / 1000 * dataset_scenario$Diff
)
```
Finally, the overall carbon intensity production for both GB and EU are reported in the following bar plot.

```{r,fig.dim = c(15,9)}
#Carbon intensity UK/EU
data <- data.frame(
  Year = rep(c("2020", '2021', '2022', '2023'), 6),
  State = c(
    rep('DK', 4),
    rep('FR', 4),
    rep('IR', 4),
    rep('UK', 4),
    rep('NL', 4),
    rep('BE', 4)
  ),
  CI = c(
    207,
    273,
    231,
    163,
    51,
    52,
    72,
    45,
    350,
    378,
    405,
    382,
    240,
    266,
    239,
    215,
    408,
    412,
    370,
    310,
    214,
    152,
    174,
    172
  )
)

data %>%
  ggplot(aes(
    x = Year,
    y = CI,
    group = State,
    fill = State
  )) +
  geom_col(
    position = position_dodge(width = 0.8),
    colour = "black",
    width = 0.65
  ) +
  geom_hline(aes(yintercept = 0)) + ylab('gCo2/KW') + ggtitle('Annual Carbon Intensity Electricity Production') +
  scale_colour_manual(values =  viridis(6),
                      aesthetics = c("colour", "fill"))
```


```{r}
knitr::knit_exit()
```

```{r}
df <- data.frame(Exports = c("Europe", "Europe", "Europe", "Europe", "Others", "Others", "Others", "Others"),
                 State = c("Montenegro", "Serbia", "North Macedonia","United Kingdom", "Montenegro", "Serbia", "North Macedonia","United Kingdom"),
                 freq = c(27,31,52,100,73,69,48,0))

df$State <- factor(df$State,
                   levels = c("United Kingdom", "North Macedonia", "Serbia", "Montenegro"),
                   ordered = TRUE)

df$Exports <- factor(df$Exports,
                     levels = c("Others","Europe"),
                     ordered = TRUE)

ggplot(df, aes(x=freq, y=State, fill=Exports))+
  geom_col(width = 0.6)+
  scale_fill_manual(name = 'Exported Electricity',values=c("blue", "red")) + xlab('%')
```


